---
title: "Predicting cross-country variability in broad interest in science differences between boys and girls with country-level gender equality."
output:
  html_document: 
    df_print: default
    toc: yes
    number_sections: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparations

## Packages

```{r message=FALSE, warning=FALSE}
library(multid)
library(lmerTest)
library(rio)
library(dplyr)
library(tibble)
library(ggpubr)
library(ggplot2)
library(MetBrewer)
library(emmeans)
library(finalfit)
source("../../custom_functions.R")
```

## Data

```{r}

coltypes<-c("text","numeric","numeric","numeric","numeric",
            "text","numeric","numeric","numeric","text",
            "text","text","text","numeric","text",
            "text","text","text")

# individual-level data
fdat<-import("../data/processed/fdat.xlsx",
             col_types=coltypes)
# country-level data
cdat<-import("../data/processed/cdat_processed.xlsx")
```

## Data exclusions and transformations

```{r}
fdat<-fdat %>%
  dplyr::select(CNT,Sex,INTBRSCI,GGGI_2015) %>%
  na.omit() %>%
  mutate(sex.c=ifelse(Sex==1,-0.5,0.5))

# standardize predictor

GGGI_2015 <- tapply(fdat$GGGI_2015,
                         fdat$CNT,
                         mean,
                         na.rm = TRUE
  )

fdat$GGGI_2015.z<-
  (fdat$GGGI_2015-mean(GGGI_2015))/
  sd(GGGI_2015)


```


# Analysis

## Reliability of the difference score

```{r}

reliab.INTBRSCI<-
  reliability_dms(
    data=fdat,
    diff_var="sex.c",var = "INTBRSCI",
    diff_var_values = c(0.5,-0.5),
    group_var = "CNT")


export(t(data.frame(reliab.INTBRSCI)),
       "../results/reliab.INTBRSCI.xlsx",
       overwrite=T)
reliab.INTBRSCI

```

## Multi-level model

### Fit model

```{r}

fit_INTBRSCI<-
  ddsc_ml(data = fdat,predictor = "GGGI_2015",
          moderator = "sex.c",moderator_values=c(0.5,-0.5),
          DV = "INTBRSCI",lvl2_unit = "CNT",re_cov_test = T,
          scaling_sd = "observed")
```

### Descriptive statistics

```{r}
export(rownames_to_column(data.frame(fit_INTBRSCI$descriptives)),
       "../results/INTBRSCI_ml_desc.xlsx",
       overwrite=T)
round(fit_INTBRSCI$descriptives,2)
round(fit_INTBRSCI$SDs,2)
```

### Variance heterogeneity test

```{r}
export(t(data.frame(fit_INTBRSCI$re_cov_test)),
       "../results/INTBRSCI_ml_var_test.xlsx",
       overwrite=T)
round(fit_INTBRSCI$re_cov_test,3)
```

### Component correlation

```{r}
export(rownames_to_column(data.frame(fit_INTBRSCI$ddsc_sem_fit$variance_test)),
       "../results/INTBRSCI_ml_comp_cor.xlsx",
       overwrite=T)
round(fit_INTBRSCI$ddsc_sem_fit$variance_test,3)
```

### Deconstructing results

```{r}
export(rownames_to_column(data.frame(fit_INTBRSCI$results)),
       "../results/INTBRSCI_ml_results.xlsx",
       overwrite=T)
round(fit_INTBRSCI$results,3)
```

### Multi-level model output

```{r}
# cross-level interaction model
summary(fit_INTBRSCI$model)

# reduced model without the predictor
summary(fit_INTBRSCI$reduced_model)
```

## Country-level path model

### Fit the model

The model is already stored within the multi-level model object. 

```{r}
fit_INTBRSCI_sem<-fit_INTBRSCI$ddsc_sem_fit
```

### Results

These are the same for both modeling techniques

```{r}
export(rownames_to_column(data.frame(fit_INTBRSCI_sem$results)),
       "../results/INTBRSCI_sem_results.xlsx",
       overwrite=T)
round(fit_INTBRSCI_sem$results,3)
```

## Plot the results

```{r}
# start with obtaining predicted values for means and differences

# refit reduced and full models with GGGI in original scale

ml_INTBRSCI_red<-
  lmer(INTBRSCI~sex.c+(sex.c|CNT),data=fdat,
       control = lmerControl(optimizer="bobyqa",
                             optCtrl=list(maxfun=2e6)))

# refit the model with raw variable

ml_INTBRSCI<-
  lmer(INTBRSCI~sex.c*GGGI_2015+(sex.c|CNT),data=fdat,
       control = lmerControl(optimizer="bobyqa",
                             optCtrl=list(maxfun=2e6)))

# point predictions as function of GGGI for components


p<-
  emmip(
    ml_INTBRSCI, 
    sex.c ~ GGGI_2015,
    at=list(sex.c = c(-0.5,0.5),
            GGGI_2015=
              seq(from=round(range(fdat$GGGI_2015,na.rm=T)[1],2),
                  to=round(range(fdat$GGGI_2015,na.rm=T)[2],2),
                  by=0.001)),
    plotit=F,CIs=T,lmerTest.limit = 1e6,disable.pbkrtest=T)

p$sex<-p$tvar
levels(p$sex)<-c("Girls","Boys")

# obtain min and max for aligned plots
min.y.comp<-min(p$LCL)
max.y.comp<-max(p$UCL)

# Boys and Girls mean distributions

p3<-coefficients(ml_INTBRSCI_red)$CNT
p3<-cbind(rbind(p3,p3),weight=rep(c(-0.5,0.5),each=nrow(p3)))
p3$xvar<-p3$`(Intercept)`+p3$sex.c*p3$weight
p3$sex<-as.factor(p3$weight)
levels(p3$sex)<-c("Girls","Boys")

# obtain min and max for aligned plots
min.y.mean.distr<-min(p3$xvar)
max.y.mean.distr<-max(p3$xvar)

# obtain the coefs for the sex-effect (difference) as function of GGGI

p2<-data.frame(
  emtrends(ml_INTBRSCI,var="sex.c",
           specs="GGGI_2015",
           at=list(#Sex = c(-0.5,0.5),
             GGGI_2015=
               seq(from=round(range(fdat$GGGI_2015,na.rm=T)[1],2),
                   to=round(range(fdat$GGGI_2015,na.rm=T)[2],2),
                   by=0.001)),
           lmerTest.limit = 1e6,disable.pbkrtest=T))

p2$yvar<-p2$sex.c.trend
p2$xvar<-p2$GGGI_2015
p2$LCL<-p2$lower.CL
p2$UCL<-p2$upper.CL

# obtain min and max for aligned plots
min.y.diff<-min(p2$LCL)
max.y.diff<-max(p2$UCL)

# difference score distribution

p4<-coefficients(ml_INTBRSCI_red)$CNT
p4$xvar=(+1)*p4$sex.c

# obtain mix and max for aligned plots

min.y.diff.distr<-min(p4$xvar)
max.y.diff.distr<-max(p4$xvar)

# define mins and maxs

min.y.pred<-
  ifelse(min.y.comp<min.y.mean.distr,min.y.comp,min.y.mean.distr)

max.y.pred<-
  ifelse(max.y.comp>max.y.mean.distr,max.y.comp,max.y.mean.distr)

min.y.narrow<-
  ifelse(min.y.diff<min.y.diff.distr,min.y.diff,min.y.diff.distr)

max.y.narrow<-
  ifelse(max.y.diff>max.y.diff.distr,max.y.diff,max.y.diff.distr)

# Figures 

# p1

# scaled simple effects to the plot

pvals<-p_coding(c(fit_INTBRSCI$results["b_21","p.value"],
                    fit_INTBRSCI$results["b_11","p.value"]))

ests<-
  round_tidy(c(fit_INTBRSCI$results["b_21","estimate"],
               fit_INTBRSCI$results["b_11","estimate"]),2)

coef1<-paste0("b21 = ",ests[1],", p = ",pvals[1])
coef2<-paste0("b11 = ",ests[2],", p = ",pvals[2])

coef_q<-round_tidy(fit_INTBRSCI$results["q_b11_b21","estimate"],2)
coef_q<-paste0("q_b = ",coef_q,", p ",
               ifelse(fit_INTBRSCI$results["interaction","p.value"]<.001,"","="),
               p_coding(fit_INTBRSCI$results["interaction","p.value"]))

coefs<-data.frame(sex=c("Girls","Boys"),
                  coef=c(coef1,coef2))

p1.INTBRSCI<-ggplot(p,aes(y=yvar,x=xvar,color=sex))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.5)+
  xlab("Global Gender Gap Index")+
  #ylim=c(2.3,3.9)+
  ylim(c(min.y.pred,max.y.pred))+
  ylab("Interest in Science")+
  scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "top",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(linewidth = 0.5, linetype = 2,
                                          colour = "gray"))+
  geom_text(data = coefs,show.legend=F,
            aes(label=coef,x=0.65,
                y=c(round(min(p$LCL),2)+0.05-0.05
                    ,round(min(p$LCL),2)-0.05),size=14,hjust="left"))+
  geom_text(inherit.aes=F,aes(x=0.65,y=round(min(p$LCL),2)-0.10,
                              label=coef_q,size=14,hjust="left"),
            show.legend=F)
p1.INTBRSCI

# prediction plot for difference score

pvals2<-p_coding(fit_INTBRSCI$results["r_xy1y2","p.value"])

ests2<-
  round_tidy(fit_INTBRSCI$results["r_xy1y2","estimate"],2)

coefs2<-paste0("r = ",ests2,
               ", p ",
               ifelse(fit_INTBRSCI$results["r_xy1y2","p.value"]<.001,"","="),
               pvals2)

p2.INTBRSCI<-ggplot(p2,aes(y=yvar,x=xvar))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.5)+
  xlab("Global Gender Gap Index")+
  ylim(c(min.y.narrow,max.y.narrow))+
  ylab("Difference in Interest in Science")+
  #scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "right",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(size = 0.5, linetype = 2,
                                          colour = "gray"))+
  #geom_text(coef2,aes(x=0.63,y=min(p2$LCL)))
  geom_text(data = data.frame(coefs2),show.legend=F,
            aes(label=coefs2,x=0.65,
                y=c(round(min(p2$LCL),2)),size=14,hjust="left"))
p2.INTBRSCI

# mean-level distributions

p3.INTBRSCI<-
  ggplot(p3, aes(x=xvar, fill=sex)) + 
  geom_density(alpha=.75) + 
  scale_fill_manual(values=met.brewer("Archambault")[c(6,2)])+
  #scale_fill_manual(values=c("turquoise3","orangered2","black")) + 
  xlab("")+
  ylab("Density")+
  ylim(c(0,6))+
  xlim(c(min.y.pred,max.y.pred))+
  theme_bw()+
  theme(legend.position = "top",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.border = element_blank(),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(size = 0.5, linetype = 2,
                                          colour = "gray"))+
  coord_flip()
p3.INTBRSCI

# distribution for mean differences

p4.INTBRSCI<-
  ggplot(p4, aes(x=xvar,fill="black")) + 
  geom_density(alpha=.75) + 
  scale_fill_manual(values="black")+
  #scale_fill_manual(values=c("turquoise3","orangered2","black")) + 
  xlab("")+
  ylab("Density")+
  ylim(c(0,6))+
  xlim(c(min.y.narrow,max.y.narrow))+
  theme_bw()+
  theme(legend.position = "none",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.border = element_blank(),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(size = 0.5, linetype = 2,
                                          colour = "gray"))+
  coord_flip()
p4.INTBRSCI

# combine component-specific predictions

p13.INTBRSCI<-
  ggarrange(p1.INTBRSCI, p3.INTBRSCI,common.legend = T,
            ncol=2, nrow=1,widths=c(4,1.4)
  )

# combine difference score predictions

p24.INTBRSCI<-
  ggarrange(p2.INTBRSCI, p4.INTBRSCI,
            ncol=2, nrow=1,widths=c(4,1.4)
  )

pall.INTBRSCI<-
  ggarrange(p13.INTBRSCI,p24.INTBRSCI,align = "hv",
            ncol=1,nrow=2,heights=c(2,1))
pall.INTBRSCI


png(filename = 
      "../results/pall.INTBRSCI.png",
    units = "cm",
    width = 21.0,height=29.7*(4/5),res = 600)
pall.INTBRSCI
dev.off()
```

# Session Information

```{r}
s<-sessionInfo()
print(s,locale=F)
```
